### 2019-09-25 MSB1015_Assignment2_SuzannetenHage.Rmd
Initial commit
Start of Assignment 2. 
First draft contains working query.

---
title: "MSB1015_Assignment 2"
author: "Suus ten Hage"
date: "25 september 2019"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
**Project Description**
In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure. This is based on data from Wiener, 1947.

** 0. Load packages**
If you haven't installed the packages on your computer yet, please uncomment the 
install.packages lines.

```{r}
# install.packages("WikidataQueryServiceR")
library("WikidataQueryServiceR")
```

** 1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all alkines with known boiling points from WikiData. To do this, I first select all chemical compounds (Q11173), their labels and their boiling points (P2102). Next, I select all compounds with a label that is in English. Finally, I select the compounds whose label ends with "ane" to select the alkanes. This results in 91 alkanes and their boiling points.

```{r}
data = query_wikidata('SELECT DISTINCT ?alkanes ?alkaneLabel ?boilingPoint      WHERE {
    ?alkanes (wdt:P31|wdt:P279) wd:Q11173;
      rdfs:label ?alkaneLabel;
      wdt:P2102 ?boilingPoint.
    FILTER((LANG(?alkaneLabel)) = "en")
    FILTER(STRENDS(?alkaneLabel, "ane"))
  }'
)
```

### 2019-09-25 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Query is changed to only include alkanes without complex structures (so only containing C's and H's). This is an edit from egonw's code. 

Added code to obtain structure descriptors.

@@
 output: html_notebook
-editor_options: 
-  chunk_output_type: inline
 ---
+
 **Project Description**
@@
 
-** 0. Load packages**
+**0. Load packages**
 If you haven't installed the packages on your computer yet, please uncomment the 
@@
 library("WikidataQueryServiceR")
+#install.packages('rJava')
+library('rJava')
+#install.packages('rcdk')
+library('rcdk')
+#install.packages('pls')
+library(pls)
+
 ```
 
-** 1. Get the data from WikiData.** 
+**1. Get the data from WikiData.** 
 In this section I use the WikidataQueryServiceR package to obtain all alkines with known boiling points from WikiData.
 To do this, I first select all chemical compounds (Q11173), their labels and their boiling points (P2102).
-Next, I select all compounds with a label that is in English.
-Finally, I select the compounds whose label ends with "ane" to select the alkanes.
-This results in 91 alkanes and their boiling points.
+Next, I select all compounds with a label that is in English and whose label ends with "ane" to select the alkanes.
+Finally, we select for those alkanes that also have their smile available in the WikiData database.
+This results in 89 alkanes and their boiling points and smiles.
 
 ```{r}
-data = query_wikidata('SELECT DISTINCT ?alkanes ?alkaneLabel ?boilingPoint      WHERE {
-    ?alkanes (wdt:P31|wdt:P279) wd:Q11173;
-      rdfs:label ?alkaneLabel;
-      wdt:P2102 ?boilingPoint.
-    FILTER((LANG(?alkaneLabel)) = "en")
-    FILTER(STRENDS(?alkaneLabel, "ane"))
-  }'
+data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
+    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
+    p:P2102 [ps:P2102 ?bp ;           
+            psv:P2102/wikibase:quantityUnit  
+            ?bpUnit         
+            ] ;
+    wdt:P233 ?smiles.
+    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
+    } '
 )
+
 ```
+
+**3.Quantify the alkane structures**
+
+```{r}
+
+#Get descriptors
+parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)
+
+descNames <- c(
+'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
+'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor'
+)
+
+descriptors <- eval.desc(parsed_smiles, descNames)
+```

### 2019-09-25 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Updated the description above the query (in last commit query and description didn't match).

@@
 **Project Description**
-In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure. This is based on data from Wiener, 1947.
+In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure.
 
@@
 **1. Get the data from WikiData.** 
-In this section I use the WikidataQueryServiceR package to obtain all alkines with known boiling points from WikiData.
-To do this, I first select all chemical compounds (Q11173), their labels and their boiling points (P2102).
-Next, I select all compounds with a label that is in English and whose label ends with "ane" to select the alkanes.
-Finally, we select for those alkanes that also have their smile available in the WikiData database.
-This results in 89 alkanes and their boiling points and smiles.
+In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData.
 
@@
 descriptors <- eval.desc(parsed_smiles, descNames)
-```

### 2019-09-25 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Added conversion of all non-kelvin temperature to kelvin.

@@
 #install.packages('pls')
-library(pls)
+library('pls')
 
@@
 descriptors <- eval.desc(parsed_smiles, descNames)
+```
+
+**4. Data pre-processing**
+Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.
+
+```{r}
+#Find all rownumbers of the values in Celcius. 
+celcius <-  which(grepl('degree Celsius', data$bpUnitLabel))
+for (i in 1:length(celcius)) {
+  #Convert the value to kelvin
+  data$bp[celcius[i]] <- data$bp[celcius[i]] + 273.15
+  #Convert the label to kelvin. 
+  data$bpUnitLabel[celcius[i]] <- 'kelvin'
+}
+
+#Find all rownumbers of the values in Fahrenheit. 
+fahrenheit <- which(grepl('degree Fahrenheit', data$bpUnitLabel))
+for (i in 1:length(fahrenheit)) {
+  #Convert the value to kelvin
+  data$bp[fahrenheit[i]] <- (data$bp[fahrenheit[i]] + 459.67) * (5/9)
+  #Convert the label to kelvin. 
+  data$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
+}
+
+```

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
updated descriptions
added null model  
added PLS
added visualizations of data and PLS results

@@
 **0. Load packages**
-If you haven't installed the packages on your computer yet, please uncomment the 
-install.packages lines.
+If you haven't installed the packages on your computer yet, please uncomment the install.packages lines before running the code below.
 
 ```{r}
+
 # install.packages("WikidataQueryServiceR")
@@
 library('rcdk')
+#install.packages('caTools')
+library('caTools')
 #install.packages('pls')
 library('pls')
+#install.packages('Metrics')
+library('Metrics')
 
@@
 **1. Get the data from WikiData.** 
-In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData.
+In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData. I get their name, boiling point, boiling point unit and smiles.
 
 ```{r}
+
 data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
@@
 
-**3.Quantify the alkane structures**
+**2. Get the descriptors of the alkane structures**
+I get descriptors for all alanes from CDK based on their smiles. These descriptors will be used later as predictors in the PLS model.
 
@@
 
-#Get descriptors
+#Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
 parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)
 
+#Determine which descriptors to get. 
 descNames <- c(
 'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
-'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor'
-)
+'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor')
 
+#Get descriptors
 descriptors <- eval.desc(parsed_smiles, descNames)
+
 ```
 
-**4. Data pre-processing**
-Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.
+**3. Data pre-processing**
+3a) Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.
 
 ```{r}
+
 #Find all rownumbers of the values in Celcius. 
@@
 ```
+
+3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
+In the following code you can easily change the size of the test data by adjusting percentage_test.
+
+```{r}
+
+#Percentage of data to go into the test-set. 
+percentage_test <- 0.25
+
+#Splitting data into a training and test set. 
+set.seed(50) 
+sample = sample.split(full_data, SplitRatio = percentage_test)
+train = subset(full_data, sample == FALSE)
+test  = subset(full_data, sample == TRUE)
+
+```
+
+**4. Null model**
+The null model is the most simple model possible: the mean of the data. This model can be used as a reference to see whether the PLS model performs better.
+
+```{r}
+
+#Create null_model
+null_model <- mean(train$data.bp)
+
+#Get error of null model
+error_null_model <- rmse(test$data.bp, null_model)
+
+```
+
+**5. Partial Least Squares**
+Here, I train the PLS model with the training data, and use the model to predict the boiling points of the test data.
+
+```{r}
+
+#Training of the PLS model. 
+PLS_model <- plsr(data.bp ~ (apol + WPATH + WPOL), 
+                  data = train, 
+                  validation = "CV"
+                  )
+
+#Predict the boiling points of the test set. 
+predicted.bp <- predict(PLS_model, test)
+
+#Calculate the error. 
+error <- RMSEP(PLS_model, newdata = test)
+
+```
+
+**6. Visualization**
+
+```{r}
+
+#Plot the predicted data vs. the original data. 
+plot(test$data.bp, predicted.bp[,,3], xlim=c(250,1000), ylim=c(250,1000))

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
I ran the same script, but this time normalized the data before training the models to see whether it improved the results. It did not.

@@
 
-data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
+data_raw = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
     ?comp wdt:P31/wdt:P279* wd:Q41581 ;
@@
 #Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
-parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)
+parsed_smiles <- parse.smiles(data_raw$smiles, kekulise=TRUE)
 
@@
 #Find all rownumbers of the values in Celcius. 
-celcius <-  which(grepl('degree Celsius', data$bpUnitLabel))
+celcius <-  which(grepl('degree Celsius', data_raw$bpUnitLabel))
 for (i in 1:length(celcius)) {
   #Convert the value to kelvin
-  data$bp[celcius[i]] <- data$bp[celcius[i]] + 273.15
+  data_raw$bp[celcius[i]] <- data_raw$bp[celcius[i]] + 273.15
   #Convert the label to kelvin. 
-  data$bpUnitLabel[celcius[i]] <- 'kelvin'
+  data_raw$bpUnitLabel[celcius[i]] <- 'kelvin'
 }
@@
 #Find all rownumbers of the values in Fahrenheit. 
-fahrenheit <- which(grepl('degree Fahrenheit', data$bpUnitLabel))
+fahrenheit <- which(grepl('degree Fahrenheit', data_raw$bpUnitLabel))
 for (i in 1:length(fahrenheit)) {
   #Convert the value to kelvin
-  data$bp[fahrenheit[i]] <- (data$bp[fahrenheit[i]] + 459.67) * (5/9)
+  data_raw$bp[fahrenheit[i]] <- (data_raw$bp[fahrenheit[i]] + 459.67) * (5/9)
   #Convert the label to kelvin. 
-  data$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
+  data_raw$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
 }
@@
 
-3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
+3b) Normalization of the data.
+
+```{r} 
+#Normalization of data
+data <- data.frame(descriptors, data_raw$bp)
+
+#Devide every column by its maximum value. 
+for (i in 1:ncol(data)) {
+  data[,i] <- data[,i] * (1/max(data[,i]))
+}
+
+```
+
+3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
 In the following code you can easily change the size of the test data by adjusting percentage_test.
@@
 set.seed(50) 
-sample = sample.split(full_data, SplitRatio = percentage_test)
-train = subset(full_data, sample == FALSE)
-test  = subset(full_data, sample == TRUE)
+sample = sample.split(data, SplitRatio = percentage_test)
+train = subset(data, sample == FALSE)
+test  = subset(data, sample == TRUE)
 
@@
 #Create null_model
-null_model <- mean(train$data.bp)
+null_model <- mean(train$data_raw.bp)
 
 #Get error of null model
-error_null_model <- rmse(test$data.bp, null_model)
+error_null_model <- rmse(test$data_raw.bp, null_model)
 
@@
 #Training of the PLS model. 
-PLS_model <- plsr(data.bp ~ (apol + WPATH + WPOL), 
+PLS_model <- plsr(data_raw.bp ~ (apol + WPATH + WPOL), 
                   data = train, 
@@
 #Plot the predicted data vs. the original data. 
-plot(test$data.bp, predicted.bp[,,3], xlim=c(250,1000), ylim=c(250,1000))
+plot(test$data_raw.bp, predicted.bp[,,3])
+
+```

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Fixed a bug: I forgot to create a data structure containing both the predictors and boiling points. This is fixed now.

@@
 parsed_smiles <- parse.smiles(data_raw$smiles, kekulise=TRUE)
-
 #Determine which descriptors to get. 
@@
 'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor')
-
 #Get descriptors
@@
 }
-
 #Find all rownumbers of the values in Fahrenheit. 
@@
 
-3b) Normalization of the data.
-
-```{r} 
-#Normalization of data
-data <- data.frame(descriptors, data_raw$bp)
-
-#Devide every column by its maximum value. 
-for (i in 1:ncol(data)) {
-  data[,i] <- data[,i] * (1/max(data[,i]))
-}
-
-```
-
-3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
+3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
 In the following code you can easily change the size of the test data by adjusting percentage_test.
@@
 ```{r}
+
+data <- data.frame(descriptors, data_raw$bp)
 
@@
 #Plot the predicted data vs. the original data. 
-plot(test$data_raw.bp, predicted.bp[,,3])
+plot(test$data_raw.bp, predicted.bp[,,3], xlim=c(350,1000), ylim=c(350,1000))
+cor(predicted.bp[,,3], test$data_raw.bp)
 

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Added two more descriptors to make predictions more precise. 
Added an error  vs. number of components plot.

@@
 parsed_smiles <- parse.smiles(data_raw$smiles, kekulise=TRUE)
+
 #Determine which descriptors to get. 
@@
 'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
-'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor')
+'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor',
+'org.openscience.cdk.qsar.descriptors.molecular.MDEDescriptor',
+'org.openscience.cdk.qsar.descriptors.molecular.FMFDescriptor'
+)
+
 #Get descriptors
@@
 #Splitting data into a training and test set. 
-set.seed(50) 
+set.seed(101) 
 sample = sample.split(data, SplitRatio = percentage_test)
@@
 #Training of the PLS model. 
-PLS_model <- plsr(data_raw.bp ~ (apol + WPATH + WPOL), 
+PLS_model <- plsr(data_raw.bp ~  ., 
                   data = train, 
@@
 ```{r} 
+#Plot the error vs. the number of components
+plot(error)
 
 #Plot the predicted data vs. the original data. 
-plot(test$data_raw.bp, predicted.bp[,,3], xlim=c(350,1000), ylim=c(350,1000))
-cor(predicted.bp[,,3], test$data_raw.bp)
+plot(test$data_raw.bp, predicted.bp[,,10], xlim=c(350,1000), ylim=c(350,1000))
+
+#Correlation between predicted and original test data. 
+corcoef <- cor(predicted.bp[,,10], test$data_raw.bp)
 

### 2019-10-12 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved package installation and project description
* In the previous version, it was necessary to manually check whether packages were installed and installing missing packages. This is now automated. 
* Documentation of section 0 (on installation) is updated . 
* The project description has been updated to include a more accurate description of the project, because it was outdated. Also, a missing reference was added.

@@
 title: "MSB1015_Assignment 2"
-author: "Suus ten Hage"
-date: "25 september 2019"
+author: "Suzanne ten Hage"
+date: "12 october 2019"
 output: html_notebook
@@
 **Project Description**
-In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure.
+Chemical properties, such as the boiling point, can be derived from the structure of a chemical compound.
+In 1947, Harry Wiener already made a correlation model to link structural features to boiling points (ref1).
+The idea to use mathematical models to predict chemical properties from compound structures has been expaned since then.
 
-**0. Load packages**
-If you haven't installed the packages on your computer yet, please uncomment the install.packages lines before running the code below.
+In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData.
+I use the smiles to get descriptors from the CDK database.
+These descriptors contain information on the structural properties of the alkanes (see section 2 for more details).
+Finally, I train a Partial Least Squares (PLS) model to predict these properties from the chemical properties of the compounds and plot the results.
+
+**0. Installation**
+The project requires several packages. The piece of code in this section checks automatically for missing packages and installs them.
+
+The rJava package requires Java to be installed.
+The code has been developed using Java version 1.8.0_191.
+A tutorial on how to install this Java Version can be found here: https://downlinko.com/download-install-jdk-8-windows.html
 
 ```{r}
+#Required packages for this code. 
+packages <- c("WikidataQueryServiceR", 
+              "rJava", 
+              "rcdk", 
+              "caTools",
+              "pls", 
+              "Metrics"
+              )
 
-# install.packages("WikidataQueryServiceR")
+#Installation of missing packages 
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+    install.packages(setdiff(packages, rownames(installed.packages()))) 
+    }
+
+#Load packages
 library("WikidataQueryServiceR")
-#install.packages('rJava')
 library('rJava')
-#install.packages('rcdk')
 library('rcdk')
-#install.packages('caTools')
 library('caTools')
-#install.packages('pls')
 library('pls')
-#install.packages('Metrics')
 library('Metrics')
@@
 
-data_raw = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
+data_wikidata = query_wikidata('SELECT DISTINCT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
     ?comp wdt:P31/wdt:P279* wd:Q41581 ;
@@
 **2. Get the descriptors of the alkane structures**
-I get descriptors for all alanes from CDK based on their smiles. These descriptors will be used later as predictors in the PLS model.
+In this section, I get descriptors for all alkanes from CDK based on their smiles. These descriptors will be used later as predictors in the PLS model.
 
@@
 #Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
-parsed_smiles <- parse.smiles(data_raw$smiles, kekulise=TRUE)
+parsed_smiles <- parse.smiles(data_wikidata$smiles, kekulise=TRUE)
 
 #Determine which descriptors to get. 
-descNames <- c(
+descriptor_names <- c(
 'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
@@
 'org.openscience.cdk.qsar.descriptors.molecular.MDEDescriptor',
-'org.openscience.cdk.qsar.descriptors.molecular.FMFDescriptor'
+'org.openscience.cdk.qsar.descriptors.molecular.FragmentComplexityDescriptor'
 )
@@
 #Get descriptors
-descriptors <- eval.desc(parsed_smiles, descNames)
+data_descriptors <- eval.desc(parsed_smiles, descriptor_names)
 
@@
 
+##CELCIUS -> KELVIN
 #Find all rownumbers of the values in Celcius. 
-celcius <-  which(grepl('degree Celsius', data_raw$bpUnitLabel))
+celcius <-  which(grepl('degree Celsius', data_wikidata$bpUnitLabel))
+
 for (i in 1:length(celcius)) {
   #Convert the value to kelvin
-  data_raw$bp[celcius[i]] <- data_raw$bp[celcius[i]] + 273.15
+  data_wikidata$bp[celcius[i]] <- data_wikidata$bp[celcius[i]] + 273.15
   #Convert the label to kelvin. 
-  data_raw$bpUnitLabel[celcius[i]] <- 'kelvin'
+  data_wikidata$bpUnitLabel[celcius[i]] <- 'kelvin'
 }
+
+##FAHRENHEIT -> KELVIN
 #Find all rownumbers of the values in Fahrenheit. 
-fahrenheit <- which(grepl('degree Fahrenheit', data_raw$bpUnitLabel))
+fahrenheit <- which(grepl('degree Fahrenheit',data_wikidata$bpUnitLabel))
+
 for (i in 1:length(fahrenheit)) {
   #Convert the value to kelvin
-  data_raw$bp[fahrenheit[i]] <- (data_raw$bp[fahrenheit[i]] + 459.67) * (5/9)
+  data_wikidata$bp[fahrenheit[i]] <- (data_wikidata$bp[fahrenheit[i]] + 459.67) * (5/9)
   #Convert the label to kelvin. 
-  data_raw$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
+  data_wikidata$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
 }
@@
 
-3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
+3b) Combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS. I remove the columns with (almost) all zeros, because 
+```{r}
+
+#Fuse the boiling points from data.raw and descriptors into one dataset. 
+data_combined <- data.frame(data_descriptors, data_wikidata$bp)
+
+#Define which columns to remove. 
+remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", "MDEN.33")
+
+#Remove all empty columns from the dataset.
+data_combined <- data_combined[ , -which(names(data_combined) %in% remove_columns)]
+
+```
+
+3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
 In the following code you can easily change the size of the test data by adjusting percentage_test.
@@
 ```{r}
-
-data <- data.frame(descriptors, data_raw$bp)
 
@@
 
+#test.try <- test[,c(1:34)];
 #Predict the boiling points of the test set. 
@@
 ```{r}
+
 #Plot the error vs. the number of components
@@
 #Plot the predicted data vs. the original data. 
-plot(test$data_raw.bp, predicted.bp[,,10], xlim=c(350,1000), ylim=c(350,1000))
+plot(test$data_raw.bp, predicted.bp[,,3], xlim=c(300,1000), ylim=c(300,1000))
+
+plot(test$data_raw.bp, rep(null_model, length(test$data_raw.bp)))
 
 #Correlation between predicted and original test data. 
-corcoef <- cor(predicted.bp[,,10], test$data_raw.bp)
+corcoef <- cor(predicted.bp[,,3], test$data_raw.bp)
 
 ```
+
+**7. References**
+ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.

### 2019-10-12 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved documentation of section 1 + added references
* The documentation of section 1, on the SPARQL query, was outdated and insufficient. This has been updated now. 
* The references for the WikiData and CDK databases were missing, these have been added.

@@
 
-In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData.
-I use the smiles to get descriptors from the CDK database.
+In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData (ref2).
+I use the smiles to get descriptors from the CDK database (ref3-6).
 These descriptors contain information on the structural properties of the alkanes (see section 2 for more details).
@@
 **1. Get the data from WikiData.** 
-In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData. I get their name, boiling point, boiling point unit and smiles.
+In this section I use the WikidataQueryServiceR package to run a SPARQL query that obtains all simple alkines with known boiling points and smiles from WikiData.
 
 ```{r}
+
+#The function query_wikidata runs the SPARQL query to obtain data from the Wikidata database.
+This returns a dataframe containing the alkane names, boiling points, units and smiles.
+It also contains the wikidata-links for compounds and units.
+There are 134 alkanes at the time of writing.
 
@@
 ```
+Description of WikiData numbers:
+* P31 = instance of
+* P279 = subclass of
+* Q41581 = alkane
+* P2102 = boiling point
+* P233 = canonical smiles
+
+The SPARQL query can easily be changed to obtain different data.
+For example, one could get the smiles of alkynes instead of alkanes by changing Q41581 into Q159226.
+Look at https://www.wikidata.org/wiki/Wikidata:Main_Page to find the right numbers.
 
@@
 ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.
+ref2: https://www.wikidata.org/wiki/Wikidata:Main_Page (12-10-2019)
+ref3: Willighagen et al. The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching. J. Cheminform. 2017; 9(3), doi:10.1186/s13321-017-0220-4
+ref4: May and Steinbeck. Efficient ring perception for the Chemistry Development Kit. J. Cheminform. 2014, doi:10.1186/1758-2946-6-3
+ref5: Steinbeck et al.
+Recent Developments of the Chemistry Development Kit (CDK) - An Open-Source Java Library for Chemo- and Bioinformatics.
+Curr.
+Pharm.
+Des.
+2006; 12(17):2111-2120, doi:10.2174/138161206777585274
+ref6: Steinbeck et al. The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics. J. Chem. Inf. Comput. Sci. 2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved documentation on section 2
Some description of which descriptors were used was missing, so this has been added.

@@
 In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData (ref2).
-I use the smiles to get descriptors from the CDK database (ref3-6).
+I use the smiles to get descriptors from the chemical development kit (CDK) database (ref3-6).
 These descriptors contain information on the structural properties of the alkanes (see section 2 for more details).
@@
 **2. Get the descriptors of the alkane structures**
-In this section, I get descriptors for all alkanes from CDK based on their smiles. These descriptors will be used later as predictors in the PLS model.
+In this section, I use four descriptors for all alkanes from CDK. I can obtain the descriptors based on the alkane smiles. These descriptors will be used later as predictors in the PLS model.
+
+*APolDescriptor:*
+This describes the sum of the atomic polarizabilities.
+
+*WienerNumbersDescriptor:*
+This gives the Wiener numbers (see ref1). It returns the Wiener Path Number and the Wiener Polarity Number.
+
+*MDEDescriptor:*
+Returns the Molecular Distance Edge.
+
+*FragmentComplexityDescriptor:*
+Returns the complexity of a system.
 
@@
 
-#Determine which descriptors to get. 
+#Determine which descriptors are of interest. To add/remove descriptors, change this vector to include the desired descriptors. 
 descriptor_names <- c(
@@
 
-#Get descriptors
+#Get descriptor values form the CDK database. 
 data_descriptors <- eval.desc(parsed_smiles, descriptor_names)
 
+#At the time of writing, this returns a dataframe of 23 variables describing 134 alkanes. 
 ```

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Section 3: error fix + improved documentation
* Section 3, splitting the data in test/training set gave an error, because an old name (data) was used as input for the functions. This has been fixed by updating the name to data_combined. 
* Made some small improvements in the documentation of section 3, to improve readability and clarity.

@@
 author: "Suzanne ten Hage"
-date: "12 october 2019"
+date: "13 october 2019"
 output: html_notebook
@@
 **3. Data pre-processing**
-3a) Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.
+Before we can use the descriptors and boiling points to train a PLS model, some data pre-processing needs to be done.
+
+**3a)** Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.
 
@@
 
-3b) Combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS. I remove the columns with (almost) all zeros, because 
-```{r}
-
-#Fuse the boiling points from data.raw and descriptors into one dataset. 
-data_combined <- data.frame(data_descriptors, data_wikidata$bp)
-
-#Define which columns to remove. 
-remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", "MDEN.33")
-
-#Remove all empty columns from the dataset.
-data_combined <- data_combined[ , -which(names(data_combined) %in% remove_columns)]
-
-```
-
-3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data.
-In the following code you can easily change the size of the test data by adjusting percentage_test.
+**3b)** Combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS.
+I remove the columns with (almost) all zeros, because these descriptors do not have much predictive value.
 
@@
 
-#Percentage of data to go into the test-set. 
+#Fuse the boiling points from data_wikidata and data_descriptors into one dataset. 
+data_combined <- data.frame(data_descriptors, data_wikidata$bp)
+
+#Define which columns to remove. To include one of these columns as a predictor, remove the column here.  
+remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", "MDEN.33")
+
+#Remove all chosen columns from the dataset.
+data_combined <- data_combined[ , -which(names(data_combined) %in% remove_columns)]
+
+#At the time of writing, this returns a dataframe of 7 variables describing 134 alkanes. 
+```
+
+**3c)** To be able to test how well the model I train performs, it is necessary to keep a part of my data aside as testing data.
+
+```{r}
+
+#Percentage of data to go into the test-set. Change this value to change the relative sizes of the training and test sets.
 percentage_test <- 0.25
 
-#Splitting data into a training and test set. 
+#Split the data into a training and test set. 
 set.seed(101) 
-sample = sample.split(data, SplitRatio = percentage_test)
-train = subset(data, sample == FALSE)
-test  = subset(data, sample == TRUE)
+sample = sample.split(data_combined, SplitRatio = percentage_test)
+train = subset(data_combined, sample == FALSE)
+test  = subset(data_combined, sample == TRUE)
 
+#At the time of writing 134 alkanes are split into a training set of 115 alkanes and test set of 19 alkanes. 
 ```

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved code section 4 + Note on section 3
* Section 4: Fixed an error that occurred because an old name was used. 
* Section 4: Added a print statement, so the code automatically prints the error of the null model. 
* Section 3: Added a note that the errors of both the null model and PLS model are dependent on the random splitting of the data into training and test set.

@@
 
+Note: The error of both the null and PLS models is dependent on the random splitting in training and test set. If you run the code, it is expected that the error differs from the error reported here.
+
 **4. Null model**
-The null model is the most simple model possible: the mean of the data. This model can be used as a reference to see whether the PLS model performs better.
+The null model is the most simple model possible; the mean of the data. This model can be used as a reference to see whether the PLS model performs better.
 
@@
 #Create null_model
-null_model <- mean(train$data_raw.bp)
+null_model <- mean(train$data_wikidata.bp)
 
 #Get error of null model
-error_null_model <- rmse(test$data_raw.bp, null_model)
+error_null_model <- rmse(test$data_wikidata.bp, null_model)
+
+#Print the error
+print(paste0("The error of the Null model is ", error_null_model))
+
+#error: 250.68 degrees Kelvin.
 

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
section 5: Changed validation to LOO, improved documentation
Section 5: 
* Fixed a small error that occurred because I was using an old name for the data (data.raw is changed into data_wikidata). 
* I realized it is better to use leave-one-out validation than cross-validation, because our dataset is small. So I changed this and added it to the documentation. 
* Added an instruction video on PLS to the documentation

@@
 The code has been developed using Java version 1.8.0_191.
-A tutorial on how to install this Java Version can be found here: https://downlinko.com/download-install-jdk-8-windows.html
+A tutorial on how to install this Java Version on windows can be found here: https://downlinko.com/download-install-jdk-8-windows.html
 
@@
 
-#Installation of missing packages 
+#Installation of missing packages.
 if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
@@
 
-#Load packages
+#Load packages.
 library("WikidataQueryServiceR")
@@
 **2. Get the descriptors of the alkane structures**
-In this section, I use four descriptors for all alkanes from CDK. I can obtain the descriptors based on the alkane smiles. These descriptors will be used later as predictors in the PLS model.
+In this section, I use four descriptors for all alkanes from CDK. I obtain the descriptors based on the alkane smiles. These descriptors are used later as predictors in the PLS model.
 
@@
 #At the time of writing, this returns a dataframe of 23 variables describing 134 alkanes.
+
 ```
@@
 **3. Data pre-processing**
-Before we can use the descriptors and boiling points to train a PLS model, some data pre-processing needs to be done.
+Before we use the descriptors and boiling points to train a PLS model, some data pre-processing needs to be done.
 
@@
 for (i in 1:length(celcius)) {
-   #Convert the value to kelvin
+   #Convert the value to kelvin.
    data_wikidata$bp[celcius[i]] <- data_wikidata$bp[celcius[i]] + 273.15
@@
 for (i in 1:length(fahrenheit)) {
-   #Convert the value to kelvin
+   #Convert the value to kelvin.
    data_wikidata$bp[fahrenheit[i]] <- (data_wikidata$bp[fahrenheit[i]] + 459.67) * (5/9)
@@
 
-**3b)** Combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS.
+**3b)** Next, I combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS.
 I remove the columns with (almost) all zeros, because these descriptors do not have much predictive value.
@@
 #At the time of writing, this returns a dataframe of 7 variables describing 134 alkanes.
+
 ```
 
-**3c)** To be able to test how well the model I train performs, it is necessary to keep a part of my data aside as testing data.
+**3c)** To be able to test how well the model I train performs, it is necessary to keep a part of my data aside as test set.
+Therefore, I split the data into a training and test set.
+I use 25% of my dataset as test set.
 
@@
 #At the time of writing 134 alkanes are split into a training set of 115 alkanes and test set of 19 alkanes.
+
 ```
@@
 
-#Create null_model
+#Create null_model.
 null_model <- mean(train$data_wikidata.bp)
 
-#Get error of null model
+#Calculate the error of the null model.
 error_null_model <- rmse(test$data_wikidata.bp, null_model)
 
-#Print the error
+#Print the error.
 print(paste0("The error of the Null model is ", error_null_model))
@@
 Here, I train the PLS model with the training data, and use the model to predict the boiling points of the test data.
+If you are unfamiliar with PLS, this is video a good introduction: https://www.youtube.com/watch?v=AxmqUKYeD-U.
+In the model I use leave-one-out (LOO) validation, because my data set is small.
 
@@
 
-#Training of the PLS model. 
-PLS_model <- plsr(data_raw.bp ~ ., 
+#Training of the PLS model, using leave-one-out validation.
+PLS_model <- plsr(data_wikidata.bp ~ ., 
                   data = train, 
-                  validation = "CV"
+                  validation = "LOO"
                   )
 
-#test.try <- test[,c(1:34)];
-#Predict the boiling points of the test set. 
-predicted.bp <- predict(PLS_model, test)
+#Calculate the error of the PLS model.
+error_PLS <- RMSEP(PLS_model, newdata = test)
 
-#Calculate the error. 
-error <- RMSEP(PLS_model, newdata = test)
+#Use the model to predict the boiling points of the test set.
+predicted_bp <- predict(PLS_model, test)
 
@@
 #Plot the predicted data vs. the original data. 
-plot(test$data_raw.bp, predicted.bp[,,3], xlim=c(300,1000), ylim=c(300,1000))
+plot(test$data_raw.bp, predicted_bp[,,3], xlim=c(300,1000), ylim=c(300,1000))
 
@@
 corcoef <- cor(predicted.bp[,,3], test$data_raw.bp)
+
+#Print the error
+print(paste0("The error of the PLS model is ", error_PLS))
 

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Added data visualization to section 3
In the data pre-processing of section 3, I added two plots to check for outliers and to check the general relationship between the size of alkanes and the boiling point.

@@
 
-**3c)** To be able to test how well the model I train performs, it is necessary to keep a part of my data aside as test set.
+**3c)** In order to check the quality of the data, it is necessary to make a few plots. First, we plot the boiling points, to see whether any of them have a very abnormal value, i.e. are an outlier.
+
+```{r}
+
+#Plot the boiling points to check for outliers. 
+plot(data_combined$data_wikidata.bp, main = "Boiling point values", ylab = "boiling point (Kelvin)", cex.main = 0.8)
+
+```
+None of the values in figure 1 have a very abnormal value. Thus, there are no outliers that need to be removed.
+
+Furthermore, one could expect that, in general, bigger alkanes have a higher boiling point.
+To test whether this relationship is also present in our dataset we plot the length of the smiles vs.
+the boiling points.
+
+```{r}
+
+#Plot length of smiles vs. boiling points to check for a correlation
+plot(stri_length(data_wikidata$smiles), data_wikidata$bp, main = "Figure 2: Length of smiles vs. boiling point", xlab = "length of smiles", ylab = "boiling point (Kelvin)", cex.main = 0.8)
+
+```
+As expected, in figure 2 we do indeed see that there is a trend towards higher boiling points for bigger smiles.
+
+**3d)** To be able to test how well the model I train performs, it is necessary to keep a part of my data aside as test set.
 Therefore, I split the data into a training and test set.

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Added documentation in section 5
Documentation of plsr, RMSEP and predict functions in section 5 was missing.

@@
 
+The plsr function trains several models with different numbers of components.
+As plsr is used here, it determines the number of components by itself (and will generally use 1 up to (descriptors - 1) components).
+It also possible to fix the number of components by using the ncomp = value argument.
+
+RSMEP will calculate the error for each model built by plsr and therefore, error_PLS will contain a vector containing the error for each number of components (see section 6 for visualization of this).
+
+The function predict will predict the boiling points based on each mode in PLS_model.
+
 **6. Visualization**

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Improve code and add documentation of section 6
* The names of the data in section 6 were not updated yet, so this has been fixed. 
* The figures are improved to include titles, labels, and legends. 
* Documentation was absent, so that is now added.

@@
               "rcdk", 
+              "stringi",
               "caTools",
@@
 library('rcdk')
+library('stringi')
 library('caTools')
@@
 #Split the data into a training and test set. 
-set.seed(101) 
+set.seed(88) 
 sample = sample.split(data_combined, SplitRatio = percentage_test)
@@
 **6. Visualization**
+**6a)** In order to identify the model with the best number of components, it is needed to plot the error vs. the number of components.
 
@@
 
-#Plot the error vs. the number of components
-plot(error)
+#Plot the error vs. the number of PLS components
+plot(error_PLS, main = "Figure 3: Error vs. number of components", xlab= "number of PLS components", ylab = "Error (Kelvin)", cex.main = 0.8)
+
+```
+
+*Conclusion:* From figure 3 we see can conclude that the error decreases with adding more components untill we have 3 PLS components.
+From there, adding more components doesn't decrease the error further.
+Therefore, from here on I use the PLS model with 3 components.
+The error is much lower than the error of the null model, indicating the PLS model performs better than the null model.
+
+**6b)** Using this PLS model with 3 components, I visualize how well it can predict boiling points based on the descriptors (described in section 2).
+To do this, I plot the predicted boiling points vs.
+the actual boiling points.
+If the prediction is perfect, we'd expect all the points to line up in a perfectly linear correlation.
+
+```{r}
 
 #Plot the predicted data vs. the original data. 
-plot(test$data_raw.bp, predicted_bp[,,3], xlim=c(300,1000), ylim=c(300,1000))
-
-plot(test$data_raw.bp, rep(null_model, length(test$data_raw.bp)))
-
-#Correlation between predicted and original test data. 
-corcoef <- cor(predicted.bp[,,3], test$data_raw.bp)
-
-#Print the error
-print(paste0("The error of the PLS model is ", error_PLS))
+plot(test$data_wikidata.bp, predicted_bp[,,3], xlim=c(300,1000), ylim=c(300,1000), main = "Figure 4: Predicted boiling points vs.
+measured boiling points", cex.main = 0.8, xlab = "measured boiling points (Kelvin)", ylab = "predicted boiling points (Kelvin)")
+lines(test$data_wikidata.bp, rep(null_model, length(test$data_wikidata.bp)), col = "red")
+lines(test$data_wikidata.bp, test$data_wikidata.bp, col = "green")
+legend("topleft", c("Null model", "Perfect model", "PLS model"), col = c("red", "green", "black"), pch = c(16), bty = 'n')
 
 ```
+
+*Conclusion*: In figure 4, we see that the PLS model (in black) performs much better in predicting the boiling points than the null model (in red).
+The PLS model comes quite close to the perfect prediction model, shown in green.
 

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Added package citations
Citations for all packages used, found using the citation() function in R.

@@
 ref6: Steinbeck et al. The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics. J. Chem. Inf. Comput. Sci. 2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y
+
+*Package references*
+* WikidataQuerySerivceR
+  Mikhail Popov (2017). WikidataQueryServiceR: API Client Library for
+  'Wikidata Query Service'. R package version 0.1.1.
+  https://CRAN.R-project.org/package=WikidataQueryServiceR
+
+*rJava
+  Simon Urbanek (2019). rJava: Low-Level R to Java Interface. R package
+  version 0.9-11. https://CRAN.R-project.org/package=rJava
+
+*rcdk
+  Guha, R. (2007). 'Chemical Informatics Functionality in R'. Journal
+  of Statistical Software 6(18)
+
+*stringi
+  Gagolewski M. and others (2019). R package stringi: Character string
+  processing facilities. http://www.gagolewski.com/software/stringi/.
+
+*caTools
+  Jarek Tuszynski (2019). caTools: Tools: moving window statistics,
+  GIF, Base64, ROC AUC, etc.. R package version 1.17.1.2.
+  https://CRAN.R-project.org/package=caTools
+
+*pls
+  Bjørn-Helge Mevik, Ron Wehrens and Kristian Hovde Liland (2019). pls:
+  Partial Least Squares and Principal Component Regression. R package
+  version 2.7-1. https://CRAN.R-project.org/package=pls
+
+*Metrics
+  Ben Hamner and Michael Frasco (2018). Metrics: Evaluation Metrics for
+  Machine Learning. R package version 0.1.4.
+  https://CRAN.R-project.org/package=Metrics

### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Small tweaks based on proofread
* Fixed some typo's
* Added student number
* Added line breaks
* Made code for all plots more tidy
* Removed one wrong comment about the error

--- 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
@@
 title: "MSB1015_Assignment 2"
-author: "Suzanne ten Hage"
-date: "13 october 2019"
+author: "Suzanne ten Hage - i6201596"
+date: "13 October 2019"
 output: html_notebook
@@
 
-**Project Description**
+**Project Description** <br/>
 Chemical properties, such as the boiling point, can be derived from the structure of a chemical compound.
@@
 
-**0. Installation**
-The project requires several packages. The piece of code in this section checks automatically for missing packages and installs them.
-
+**0. Installation** <br/>
+The project requires several packages.
+The piece of code in this section checks automatically for missing packages and installs them.
 The rJava package requires Java to be installed.
 The code has been developed using Java version 1.8.0_191.
-A tutorial on how to install this Java Version on windows can be found here: https://downlinko.com/download-install-jdk-8-windows.html
+A tutorial on how to install this Java Version on windows can be found [here](https://downlinko.com/download-install-jdk-8-windows.html).
 
-```{r}
+```{r include = TRUE}
 #Required packages for this code. 
@@
 ```
-
-**1. Get the data from WikiData.** 
+**1. Get the data from WikiData.** <br/>
 In this section I use the WikidataQueryServiceR package to run a SPARQL query that obtains all simple alkines with known boiling points and smiles from WikiData.
 
-```{r}
+```{r include = TRUE}
 
@@
 ```
-Description of WikiData numbers:
-* P31 = instance of
-* P279 = subclass of
-* Q41581 = alkane
-* P2102 = boiling point
-* P233 = canonical smiles
+
+Description of WikiData numbers: <br/>
+* P31 = instance of <br/>
+* P279 = subclass of <br/>
+* Q41581 = alkane <br/>
+* P2102 = boiling point <br/>
+* P233 = canonical smiles <br/>
 
@@
 For example, one could get the smiles of alkynes instead of alkanes by changing Q41581 into Q159226.
-Look at https://www.wikidata.org/wiki/Wikidata:Main_Page to find the right numbers.
+The numbers can be found [here](https://www.wikidata.org/wiki/Wikidata:Main_Page).
 
-**2. Get the descriptors of the alkane structures**
+**2. Get the descriptors of the alkane structures** <br/>
 In this section, I use four descriptors for all alkanes from CDK. I obtain the descriptors based on the alkane smiles. These descriptors are used later as predictors in the PLS model.
 
-*APolDescriptor:*
+*APolDescriptor:* <br/>
 This describes the sum of the atomic polarizabilities.
 
-*WienerNumbersDescriptor:*
+*WienerNumbersDescriptor:* <br/>
 This gives the Wiener numbers (see ref1). It returns the Wiener Path Number and the Wiener Polarity Number.
 
-*MDEDescriptor:*
+*MDEDescriptor:* <br/>
 Returns the Molecular Distance Edge.
 
-*FragmentComplexityDescriptor:*
+*FragmentComplexityDescriptor:* <br/>
 Returns the complexity of a system.
 
-```{r}
+```{r include = TRUE}
 
@@
 
-**3. Data pre-processing**
+**3. Data pre-processing** <br/>
 Before we use the descriptors and boiling points to train a PLS model, some data pre-processing needs to be done.
@@
 
-```{r}
+```{r include = TRUE}
 
@@
 #Define which columns to remove. To include one of these columns as a predictor, remove the column here.  
-remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", "MDEN.33")
+remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", 
+                     "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", 
+                     "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", 
+                     "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", 
+                     "MDEN.33"
+                    )
 
@@
 
-```{r}
+```{r include = TRUE}
 
 #Plot the boiling points to check for outliers. 
-plot(data_combined$data_wikidata.bp, main = "Boiling point values", ylab = "boiling point (Kelvin)", cex.main = 0.8)
+plot(data_combined$data_wikidata.bp, 
+      main = "Boiling point values", 
+      ylab = "boiling point (Kelvin)", 
+      cex.main = 0.8
+     )
 
 ```
-None of the values in figure 1 have a very abnormal value. Thus, there are no outliers that need to be removed.
 
+None of the values in figure 1 have a very abnormal value.
+Thus, there are no outliers that need to be removed.
 Furthermore, one could expect that, in general, bigger alkanes have a higher boiling point.
@@
 
-```{r}
+```{r include = TRUE}
 
 #Plot length of smiles vs. boiling points to check for a correlation
-plot(stri_length(data_wikidata$smiles), data_wikidata$bp, main = "Figure 2: Length of smiles vs. boiling point", xlab = "length of smiles", ylab = "boiling point (Kelvin)", cex.main = 0.8)
+plot(stri_length(data_wikidata$smiles), 
+      data_wikidata$bp, 
+      main = "Figure 2: Length of smiles vs. boiling point", 
+      xlab = "length of smiles", 
+      ylab = "boiling point (Kelvin)", 
+      cex.main = 0.8
+     )
 
@@
 
-**4. Null model**
+**4. Null model** <br/>
 The null model is the most simple model possible; the mean of the data. This model can be used as a reference to see whether the PLS model performs better.
 
-```{r}
+```{r include = TRUE}
 
@@
 
-#error: 250.68 degrees Kelvin.
-
 ```
 
-**5. Partial Least Squares**
+**5. Partial Least Squares** <br/>
 Here, I train the PLS model with the training data, and use the model to predict the boiling points of the test data.
-If you are unfamiliar with PLS, this is video a good introduction: https://www.youtube.com/watch?v=AxmqUKYeD-U.
+If you are unfamiliar with PLS, this is [video](https://www.youtube.com/watch?v=AxmqUKYeD-U) a good introduction.
 In the model I use leave-one-out (LOO) validation, because my data set is small.
 
-```{r}
+```{r include = TRUE}
 
@@
 
-**6. Visualization**
+**6. Results** <br/>
 **6a)** In order to identify the model with the best number of components, it is needed to plot the error vs. the number of components.
 
-```{r}
+```{r include = TRUE}
 
 #Plot the error vs. the number of PLS components
-plot(error_PLS, main = "Figure 3: Error vs. number of components", xlab= "number of PLS components", ylab = "Error (Kelvin)", cex.main = 0.8)
+plot(error_PLS, 
+      main = "Figure 3: Error vs. number of components", 
+      xlab= "number of PLS components", 
+      ylab = "Error (Kelvin)", 
+      cex.main = 0.8
+     )
 
@@
 
-*Conclusion:* From figure 3 we see can conclude that the error decreases with adding more components untill we have 3 PLS components.
+*Conclusion:* <br/> 
+From figure 3 we see can conclude that the error decreases with adding more components untill we have 3 PLS components.
 From there, adding more components doesn't decrease the error further.
@@
 
-```{r}
+```{r include = TRUE, message = FALSE, warning = FALSE}
 
 #Plot the predicted data vs. the original data. 
-plot(test$data_wikidata.bp, predicted_bp[,,3], xlim=c(300,1000), ylim=c(300,1000), main = "Figure 4: Predicted boiling points vs.
-measured boiling points", cex.main = 0.8, xlab = "measured boiling points (Kelvin)", ylab = "predicted boiling points (Kelvin)")
-lines(test$data_wikidata.bp, rep(null_model, length(test$data_wikidata.bp)), col = "red")
-lines(test$data_wikidata.bp, test$data_wikidata.bp, col = "green")
-legend("topleft", c("Null model", "Perfect model", "PLS model"), col = c("red", "green", "black"), pch = c(16), bty = 'n')
+plot(test$data_wikidata.bp, 
+      predicted_bp[,,3], 
+      xlim=c(300,1000), 
+      ylim=c(300,1000), 
+      main = "Figure 4: Predicted boiling points vs. measured boiling points", 
+      cex.main = 0.8, 
+      xlab = "measured boiling points (Kelvin)", 
+      ylab = "predicted boiling points (Kelvin)"
+     )
+#Add null model to the plot
+lines(test$data_wikidata.bp, 
+       rep(null_model, 
+           length(test$data_wikidata.bp)
+          ), 
+       col = "red"
+      )
+#Add the perfect prediction to the plot
+lines(test$data_wikidata.bp, 
+       test$data_wikidata.bp, 
+       col = "green"
+      )
+#Add legend
+legend("topleft", 
+       c("Null model", "Perfect model", "PLS model"), 
+       col = c("red", "green", "black"), 
+       pch = c(16), 
+       bty = 'n'
+      )
 
@@
 
-*Conclusion*: In figure 4, we see that the PLS model (in black) performs much better in predicting the boiling points than the null model (in red).
+*Conclusion:* <br/>
+In figure 4, we see that the PLS model (in black) performs much better in predicting the boiling points than the null model (in red).
 The PLS model comes quite close to the perfect prediction model, shown in green.
 
-**7. References**
-ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.
-ref2: https://www.wikidata.org/wiki/Wikidata:Main_Page (12-10-2019)
-ref3: Willighagen et al. The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching. J. Cheminform. 2017; 9(3), doi:10.1186/s13321-017-0220-4
-ref4: May and Steinbeck. Efficient ring perception for the Chemistry Development Kit. J. Cheminform. 2014, doi:10.1186/1758-2946-6-3
+**7. References** <br/>
+ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20. <br/>
+ref2: https://www.wikidata.org/wiki/Wikidata:Main_Page (12-10-2019) <br/>
+ref3: Willighagen et al.
+The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching.
+J.
+Cheminform.
+2017; 9(3), doi:10.1186/s13321-017-0220-4 <br/>
+ref4: May and Steinbeck. Efficient ring perception for the Chemistry Development Kit. J. Cheminform. 2014, doi:10.1186/1758-2946-6-3 <br/>
 ref5: Steinbeck et al.
@@
 Des.
-2006; 12(17):2111-2120, doi:10.2174/138161206777585274
-ref6: Steinbeck et al. The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics. J. Chem. Inf. Comput. Sci. 2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y
+2006; 12(17):2111-2120, doi:10.2174/138161206777585274 <br/>
+ref6: Steinbeck et al.
+The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics.
+J.
+Chem.
+Inf.
+Comput.
+Sci.
+2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y <br/>
+
+*Package references* <br/>
+* [WikidataQuerySerivceR](https://CRAN.R-project.org/package=WikidataQueryServiceR) <br/>
+  Mikhail Popov (2017). WikidataQueryServiceR: API Client Library for
+  'Wikidata Query Service'. R package version 0.1.1. <br/>
+
+*[rJava](https://CRAN.R-project.org/package=rJava) <br/>
+  Simon Urbanek (2019). rJava: Low-Level R to Java Interface. R package
+  version 0.9-11.  <br/>
+
+*[rcdk](https://cran.r-project.org/web/packages/rcdk/index.html) <br/>
+  Guha, R. (2007). 'Chemical Informatics Functionality in R'. Journal
+  of Statistical Software 6(18) <br/>
+
+*[stringi](http://www.gagolewski.com/software/stringi/.) <br/>
+  Gagolewski M. and others (2019). R package stringi: Character string
+  processing facilities.
+
+*[caTools](https://CRAN.R-project.org/package=caTools) <br/>
+  Jarek Tuszynski (2019). caTools: Tools: moving window statistics,
+  GIF, Base64, ROC AUC, etc.. R package version 1.17.1.2 <br/>
+
+*[pls](https://CRAN.R-project.org/package=pls) <br/>
+  Bjørn-Helge Mevik, Ron Wehrens and Kristian Hovde Liland (2019). pls:
+  Partial Least Squares and Principal Component Regression. R package
+  version 2.7-1.  <br/>
+
+*[Metrics](https://CRAN.R-project.org/package=Metrics) <br/>
+  Ben Hamner and Michael Frasco (2018). Metrics: Evaluation Metrics for
+  Machine Learning. R package version 0.1.4. <br/>
